\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[left=2cm,right=2cm,bottom=3cm]{geometry}
\usepackage{fancyhdr}
\usepackage{sectsty}
\usepackage{hyperref}

% \parindent 0in
\pagestyle{fancy}
\lhead{}
\chead{}
\rhead{\bf {Jesse Wheeler}}
\lfoot{}
\cfoot {\thepage}
\rfoot{}
\title{JesseWheelerCV}
\author{jeswheel}
\date{January 2023}

\sectionfont{
   \Large
   \sectionrule{0pt}{0pt}{-4pt}{1.5pt}
}

\subsectionfont{
   \noindent \hspace{1pt} \large
}


\newcommand {\myheader}[1]{{\bf {#1}}}

\newenvironment {interests}
                {
                 \begin{list}{}
                 {\setlength{\labelwidth}{0mm}
                  %\setlength{\leftmargin}{0em}
                  \setlength{\leftmargin}{8mm}
                  \setlength{\itemindent}{-3mm}
                  \setlength{\labelsep}{0mm}
                  \setlength{\parsep}{0.1 ex}
%                  \setlength{\itemsep}{0.25 cm}
%                  \setlength{\itemsep}{0.23 cm}
                  \setlength{\itemsep}{1pt}
      \setlength{\topsep}{0.15cm}}} %space between title and 1st item
   {\end{list}}

\newenvironment {reflist}
                {
                 \begin{list}{}
                 {\setlength{\labelwidth}{0mm}
                  %\setlength{\leftmargin}{0em}
                  \setlength{\leftmargin}{8mm}
                  \setlength{\itemindent}{-3mm}
                  \setlength{\labelsep}{0mm}
                  \setlength{\parsep}{0.1 ex}
%                  \setlength{\itemsep}{0.25 cm}
%                  \setlength{\itemsep}{0.23 cm}
                  \setlength{\itemsep}{0.1cm}
      \setlength{\topsep}{0.15cm}}} %space between title and 1st item
   {\end{list}}

\newenvironment {teachinglsit}
                {
                 \begin{list}{}
                 {\setlength{\labelwidth}{0mm}
                  %\setlength{\leftmargin}{0em}
                  \setlength{\leftmargin}{5mm}
                  % \setlength{\itemindent}{0mm}
                  \setlength{\labelsep}{0mm}
                  \setlength{\parsep}{0.1 ex}
%                  \setlength{\itemsep}{0.25 cm}
%                  \setlength{\itemsep}{0.23 cm}
                  \setlength{\itemsep}{1pt}
      \setlength{\topsep}{0.15cm}}} %space between title and 1st item
   {\end{list}}

% % education entry
% \newcommand{\education}[4]{

% % name of the studies
%   \noindent  \textbf{#1}
%   % at the right the duration
%     \hfill
%     \framebox{% duration inside a frame box
%     \parbox{6em}{%
%     \centering\textbf{#2}}} \par
%  % new paragraph with the school in italics
%      \noindent \textit{#3} \par
%      % description with no hanging and in smaller text
%      \vspace*{0.5em}
%      \noindent\hangindent=2em\hangafter=0 \small #4
% %back to normal size
% \normalsize \par}

\begin{document}

\thispagestyle{empty}  % No number or header on 1st page

\rule{0mm}{1mm}
\vspace{-20mm}

%\hfill{\small \today}

\vspace{1mm}

%\begin{center}
%\hfill
\rule{0mm}{1mm}\hspace{6cm}{\large {\bf JESSE WHEELER}}
%\hfill
%\end{center}
%\vspace{1mm}

%\vspace{1mm}

\begin{tabbing}
Idaho State Universityxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill
Assistant Professor \>
{\url{https://jeswheel.github.io/}}\\
Department of Mathematics and Statistics \>  % phone: (734) 615-3332
Email: {jessewheeler@isu.edu}
\\
Idaho State University\> % fax: (734) 763-4676
\\
Pocatello, ID 83209 \>
\end{tabbing}

\section*{Education}

{\bf PhD in Statistics} \hfill {\textit{2020--2025}}

University of Michigan, Ann Arbor, MI.

Thesis Title: ``Innovations in Likelihood-Based Inference for State Space Models"

Thesis Advisor: Edward Ionides

\vspace{2mm}

\noindent{\bf B.S. in Mathematics, Statistics, Minor in Computer Science } \hfill  {\textit{2016--2020}}

Utah State University (USU), Logan, UT.

Graduated as valedictorian of the College of Science, class of 2020

\section*{Research}
\subsection*{Interests}
\begin{interests}
    \item Time Series Analysis
    \item Mechanistic Models
    \item Statistical Practice and Data Science
    \item Modeling Dynamic Ecological Systems

\end{interests}

\subsection*{Peer Reviewed Publications}

\begin{reflist}

    \item Wheeler, J., Ionides, E. L. (2025+). Revisiting Inference for ARMA Models: Improved Fits and Superior Confidence Intervals. {\it arXiv:2310.01198}. {\it PLOS One}, to appear.

    \item Bret\'o, C., Wheeler, J., King, A., Ionides, E. L. (2025). panelPomp: Analysis of Panel Data via Partially Observed Markov Processes in R. {\it arXiv:2410.07934}. {\it The R Journal}, to appear.

    \item Ionides, E. L., Ning, N. and Wheeler, J. (2024). An iterated block particle filter for inference on coupled dynamic systems with shared and unit-specific parameters. {\it Statistica Sinica}, 34, 1241-1262. doi:10.5705/ss.202022.0188.

    \item Wheeler, J., Rosengart, A., Jiang, Z., Tan, K., Truetle, N., Ionides, E. (2024). Informing policy via dynamic models: Cholera in Haiti. {\it PLOS Computational Biology}, 20(4), e1012032. \\ doi:10.1371/journal.pcbi.1012032

    \item Wagstaff, J., Bean, B., Wheeler, J., Maguire, M., Sun, Y. (2024). Adaptive Mapping of Design Ground Snow Loads in the Conterminous United States. {\it Journal of Structural Engineering}, 150(1), 04023193. doi:10.1061/JSENDH.STENG-12396

    \item Wheeler, J., Bean, B., Maguire, M. (2022). Creating a Universal Depth-to-Load Conversion Technique for the Conterminous United States Using Random Forests. {\it Journal of Cold Regions Engineering, 36(1), 04021019}. doi:10.1061/(ASCE)CR.1943-5495.0000270

    \item White, T., Wheeler, J., Lindstrom, C., Christensen, R., Moon, K. (2021). GPS-Denied Navigation Using SAR Images and Neural Networks. {\it ICASSP 2021 - 2021 IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}. Toronto, ON, Canada, 2021, pp. 2395-2399, doi:10.1109/ICASSP39728.2021.9414421.

\end{reflist}

\subsection*{Non-Refereed Publications}

\begin{reflist}

    \item Bean, B., Maguire, M., Sun, Y., Wagstaff, J., Al-Rubaye, S., Wheeler, J., Jarman, S., Rogers, M. (2021). The 2020 National Snow Load Study. {\it Utah State University}. doi:10.26077/200k-pr86.

    \item Ionides, E. L., Wheeler, J. (2024). Review 2: ``Efficacy, Public Health Impact and Optimal Use of the Takeda Dengue Vaccine." Rapid Reviews Infectious Diseases. doi:10.1162/2e3983f5.1f0cb1f4.

\end{reflist}

\subsection*{Working Papers}

\begin{reflist}


    \item Wheeler, J., Abkemeier, A. A., Ionides, E. L. (2025). Iterating Marginalized Bayes Maps for Likelihood Maximization. {\it In preperation for the Journal of the American Statistical Association}.


    \item Yang, B., Wheeler, J., King, A., Duffy, M., Ionides, E. L. (2024). Mechanistic Models for Panel Data: Analysis of Ecological Experiments with Four Interacting Species. {\it Submitted to the Journal of the American Statistical Association, Case Studies}.

\end{reflist}

\subsection*{Conference Presentations}

\begin{reflist}

    \item \textbf{2024}\, Modelling and inference for pandemic preparedness. Isaac Newton Institute, Cambridge, England.
    % \item Wheeler, J., Rosengart, A., Jiang, Z., Tan, K., Treutle, N., Ionides., E. (2023). Informing Policy Via Dynamic Models: Cholera in Haiti. Contributed Paper. {\it Modelling and inference for pandemic preparedness – a focussed workshop, Isaac Newton Institute, Cambridge, England}.

    \item \textbf{2024}\, Michigan Student Symposium for Interdisciplinary Statistical Sciences. Ann Arbor, MI.
    % \item Wheeler, J., Ionides, E. (2024). arima2: An R package for Likelihood Based Inference for ARIMA Modeling. {\it MSSISS 2024, Ann Arbor, MI.}

    \item \textbf{2023}\, Bayes Comp 2023, Satellite Event: Bayesian Inference of Epidemics. Levi, Finland.
    % \item Wheeler, J., Rosengart, A., Jiang, Z., Tan, K., Treutle, N., Ionides., E. (2023). Informing Policy Via Dynamic Models: Cholera in Haiti. Contributed Paper. {\it Bayes Comp 2023, Satellite Event: Bayesian Inference of Epidemics, Levi, Finland}.

    \item \textbf{2023}\, Models of Infectious Disease Agent Study (MIDAS) Network Annual Meeting. Atlanta, GA.

    \item \textbf{2022}\, JSM 2022. Washington D.C.
    % \item Wheeler, J., Rosengart, A., Jiang, J., Ionides, E. (2022). Eliminating Cholera in Haiti: Combining Dynamic Models with Data to Inform Vaccination Policy. Contributed Paper. {\it JSM 2022, Washington D.C.}.

     \item \textbf{2020}\, National Conference on Undergraduate Research. Online (COVID-19).

    \item \textbf{2020}\, USU Student Research Symposium. Logan, UT.
    % \item Wheeler, J., Bean, B. (2020). Estimating Snow Water Equivalence Using Easily Obtainable Climate Variables. USU Student Research Symposium.

    \item \textbf{2020}\, Utah Conference on Undergraduate Research. Logan, UT.
    % \item Wheeler, J., Lindstrom, C., Christensen, R., Moon, K. (2020). Poster: GPS-Denied Navigation Using SAR Images and Neural Networks. {\it Utah Conference on Undergraduate Research, Logan, UT}.

    % \item Christensen, D., Villanueva, I., Wheeler, J., Vicioso, P., Husman, J., Lampkins, S., Youmans, K. (2019). Exploring Potential Relationships Between Self-Efficacy, Performance, and Electrodermal Activity in Engineering Exams. {\it American Educational Research Association.} April 5-9. Toronto, Canada.

\end{reflist}

\subsection*{Awards}

\noindent\hspace{5mm}\textbf{StatsForward Fellowship}, The American Statistical Association. \hfill{\textit{September 2025}}

\vspace{2mm}
\noindent\hspace{5mm}\textbf{Best Oral Presentation}, MSSIS 2024, University of Michigan. \hfill {\textit{2024}}

\hspace{5mm}{\it Award amount: $\$200$}


\vspace{2mm}
\noindent\hspace{5mm}\textbf{Rackham Graduate Student Research Grant}, University of Michigan \hfill {\textit{2024}}

\hspace{5mm}{\it Award amount: $\$2500$}

\vspace{2mm}
\noindent\hspace{5mm}\textbf{Honorable Mention} NSF GRFP \hfill {\textit{2022}}

\vspace{2mm}
\noindent\hspace{5mm}\textbf{Rackham Merit Fellowship}, University of Michigan \hfill {\textit{2020--2025}}

\hspace{5mm}{\it Award amount: Three years of PhD student funding.}

\vspace{2mm}
\noindent\hspace{5mm}\textbf{Valedictorian}, USU College of Science \hfill {\textit{2020}}

\vspace{2mm}
\noindent\hspace{5mm}\textbf{URCO Grant}, USU \hfill  {\textit{2020}}

\hspace{5mm}{\it Award amount: $\$1000$}

\section*{Teaching}

\noindent \textbf{University of Michigan} \hfill  {\textit{2020--Present}}

\begin{teachinglsit}
    \item \textbf{Stats 604}, {\it Lab Instructor} \hfill (Statistical Practice)
    \item \textbf{Stats 531}, {\it Office hours / Grading} \hfill (Modeling and Analysis of Time Series Data)
    \item \textbf{Datasci 415}, {\it Lab Instructor} \hfill (Introduction to Statistical Learning)
    \item \textbf{Stats 306}, {\it Lab Instructor} \hfill (Introduction to Statistical Computing)
    \item \textbf{Summer Math Boot Camp} \hfill (Calculus, Linear Algebra, Probability, Computing)
    \item \textbf{Workshop} {\it Developing a personal academic website} \hfill (Student Seminar Series)
    \item \textbf{Tutor}, {\it Master's level Rackham Merit Fellows} \hfill (Probability and Regression)

\end{teachinglsit}

\vspace{2mm}
\noindent \textbf{SISMID (2025)}. Instructor for a short course on {\it Simulation based inference for Epidemiological Dynamics} at the Summer Institute in Statistics Modeling in Infectious Diseases (SISMID). Emory University, Atlanta, Georgia.

\vspace{2mm}
\noindent \textbf{SISMID (2022)}. Instructor for a short course on {\it Simulation based inference for Epidemiological Dynamics} at the Summer Institute in Statistics Modeling in Infectious Diseases (SISMID). University of Washington, Seattle.

\vspace{2mm}
\noindent \textbf{Utah State University} \hfill  {\textit{2017--2020}}

\begin{teachinglsit}
    \item Math 0995, {\it Recitation Leader} (Remedial Algebra)
    \item Math 1210, {\it Recitation Leader} (Introductory Calculus)
    \item Math 1220 {\it Recitation Leader} (Calculus II, sequences and series)
    \item Stats 1040, {\it Recitation Leader}, (Introductory Statistics, non-Calculus based)
\end{teachinglsit}

\noindent \textbf{Undergraduate Research Projects Mentored}

\begin{reflist}

  \item Peter Yang. Applications of Partially Observed Markov Processes with phylogenetic data. Honors thesis, 2025.
  \item Weizhe Sun. Model Based Inference of Stochastic Volatility via Iterated Filtering. Honors thesis, 2024.
  \item Zuyuan Han. Signature Methods in Variance Swap Pricing. Honors thesis, 2023.
  \item Bo Yang. Analysis of Panel Data via Mechanistic Models in a PanelPOMP Framework. Honors thesis, 2023.
  \item Kevin Tan and Noah Treutle. On the Transmissibility of Cholera During the 2010--2019 Haiti Cholera Epidemic. Research project, 2022.

\end{reflist}

\subsection*{Awards}

\noindent\hspace{5mm}\textbf{Outstanding Graduate Student Instructor Team Award.} \hfill {\textit{2025}}

\hspace{5mm}\textbf{Outstanding Undergraduate Recitation Leader} \hfill  {\textit{2019}}

\noindent\hspace{5mm}{\it USU Mathematics and Statistics Department}

\section*{Service}

\noindent \textbf{Computing Club Committee Member}, UM Statistics Department \hfill  {\textit{2022--Present}}\\
\noindent \textbf{Committee Chair} \hfill {\it 2023--Present}

\vspace{2mm}
\noindent \textbf{Peer Reviews}: PLOS Computation Biology, Nature Communications, Rapid Reviews Infectious Diseases, Bulletin of Mathematical Biology

\vspace{2mm}
\noindent \textbf{President, USU Data Science Club} \hfill  {\textit{2019--2020}}

\section*{Software}

\noindent Author and maintainer for the R package \textbf{arima2}, available on CRAN. As of October 1, 2024, this package has been downloaded more times than $79.7\%$ of all packages on CRAN since it was first published (October 5, 2023).

\vspace{2mm}
\noindent Primary contributor and current maintainer of the R package \textbf{panelPomp}, available on CRAN.

\vspace{2mm}
\noindent Core developer of the Python package \textbf{pypomp}, available on PyPI.

\vspace{2mm}
\noindent Contributor to open source R packages \textbf{pomp} and \textbf{spatPomp}, both available on CRAN.

\section*{Consulting}

\noindent Campfire Interactive, time series modeling. \hfill  {\textit{Nov 2024--Jan 2025}}

\end{document}
